A gossip-based protocol for exchanging immutable data in dynamic distributed environments

### 2011-07-18 papers/paper1/paper.tex
Workin' on the paper.

\title{A protocol for implementing read-only databases in dynamic distributed environments}
\author{Adam Drescher \and Magdalena Cassel}
\date{}

@@

We want to establish relationships between configuration records, which means we need to be able to download configuration records.

@@

Files are an abstraction of the configuration record.

\paragraph {Files.}
The main concept in MFTP is that of a \emph{file}.

A file is a sequence of bytes with a length and a user-defined type.
Files are divided into pieces of equal size called \emph{fragments}.
Each file is uniquely identified by a \emph{file ID} that contains the file's type, length, and a hash of the file's contents.

% TODO:  We need to move this.
% Given two files $f_1$ and $f_2$, one can define a matching predicate $\mu$.

\paragraph {Tasks.}
The two main tasks in MFTP are acquiring a file and establishing a relationship between two files.
The process of acquiring a file involves sending a request for the file's fragments and receiving fragments sent in response.
MFTP is a peer-to-peer protocol because each participant can act as both a client and a server.
A peer can send out any fragments for which it gets a request.
Fragments are periodically broadcast, even in the absence of requests, to indicate what files exist in the network.
A peer can use requests to poll for fragments of a file which can span from individual fragments to the whole file.
A peer which has successfully downloaded any number of fragments of a file can then serve those fragments.

Matching is a two-step process.
The first step applies a matching function $\iota$ to two file IDs to determine whether a match is possible between the two files.  
The second step applies a matching function $\mu$ to the two files to determine whether there is a match.
If $\mu$($f_1$,$f_2$) is true, $\iota$($fid_1$,$fid_2$) must be true.
However, if $\iota$($fid_1$,$fid_2$) is true, $\mu$($f_1$,$f_2$) does \emph{not} have to be true.

\paragraph{Messages.}
File transfer takes place via \emph{messages}.
There are three types of messages in the protocol.
The first message type is a \emph{request}.
A request contains a file ID and a set of fragment indices.
We implemented requests as the file ID and a set of contiguous regions.  
However, one can envision the request to contain anything that defines a set.
The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.

>>>>

\title{A gossip-based protocol for exchanging immutable data in dynamic distributed environments}
\author{Adam Drescher \and Magdalena Cassel}
\date{}

@@

The two main tasks in configuration record processing are acquiring records from the network and establishing a relationship between two records.
We adopt the view that configuration records can be serialized and stored in a file.
The task of acquiring a record is then equivalent to downloading a file from the network.
Once a configuration record has been downloaded, its content can then be compared to configuration records that are stored locally.

A \emph{file} in MFTP is a finite sequence of bytes with an associated type.
Each file is uniquely identified by its \emph{file ID} which contains the file's type, length, and a hash of the file's contents.
The type is defined by the user and is used to limit the potential relationships among files.
We assume the existence of a user-defined predicate $\gamma(fid_1, fid_2)$ that returns true if there is potentially a relationship between the two files.
The predicate $\gamma$ is symmetric, i.e., $\gamma(fid_1, fid_2) = \gamma(fid_2, fid_1)$ and non-reflexive, i.e., $\gamma(fid_1, fid_1)$ is false.
Note that $\gamma$ is only defined using file IDs and can therefore only check file types and lengths.
We also assume the existence of a user-defined predicate $\mu(file_1, file_2)$ that returns true if a relationship between $file_1$ and $file_2$ exists.
Like $\gamma$, the predicate $\mu$ is symmetric and non-reflexive.
The predicate $\mu$ implies $\gamma$, i.e., $\mu(file_1, file_2) \to \gamma(fid(file_1), fid(file_2))$.

MFTP assumes that all communication takes place by sending fixed-size messages over a broadcast or multicast channel.
This allows MFTP to take advantage of gossip since file contents and matching relationships can be communicated efficiently to multiple interested parties.
To enable the sending of file contents using fixed-size messaging, we divide a file into pieces of equal size called \emph{fragments}.

The MFTP protocol contains three types of messages: fragments, requests, and matches.
Each message in MFTP is prefixed with a header indicating its type.
A \emph{fragment message} consists of a file ID, a fragment index, and the data comprising the indicated fragment.
A \emph{request message} consists of a file ID and a set of fragment indices.
For example, our implementation uses a set of spans where each span requests a contiguous set of fragments.
However, one can envision the request to contain more advanced structures, e.g., functions, that can define arbitrary sets.
A \emph{match message} consist of a file ID and a set of matching file IDs for which $\gamma$ is true.

Two files with the same content and type will have the same file ID.
With a strong enough hashing mechanism, we can assume two files are equal if their file IDs are the same.
Since the identity of a file is completely determined by its content, files are immutable and can be freely replicated.
Peers in the protocol are free to cache any fragment and respond to any requests for which they have a fragment.
The process of acquiring a file consists of repeatedly sending requests and receiving fragments from any peers that chose to respond until all of the fragments have been acquired.

Learning relationships can be accomplished in one of two ways.
First, peers in MFTP periodically send a fragment to announce the presence of the file in the network.
Assume that a peer serving $file_1$ receives an announcement for $file_2$.
The peer checks $\gamma(fid(file_1), fid(file_2))$ to determine whether a relationship is possible.
If a relationship is possible, the peer acquires the file and checks $\mu(file_1, file_2)$ to determine whether the relationship exists.
Peers periodically send match messages to indicate which files have a relationship with the file they serve.
The second way relationships can be learned is by eavesdropping on the match messages sent by other peers.

%% \paragraph {Files.}
%% The main concept in MFTP is that of a \emph{file}.
%% A file is a sequence of bytes with a length and a user-defined type.
%% Files are divided into pieces of equal size called \emph{fragments}.
%% Each file is uniquely identified by a \emph{file ID} that contains the file's type, length, and a hash of the file's contents.

%% \paragraph{Messages.}
%% File transfer takes place via \emph{messages}.
%% There are three types of messages in the protocol.
%% The first message type is a \emph{request}.
%% A request contains a file ID and a set of fragment indices.
%% We implemented requests as the file ID and a set of contiguous regions.  
%% However, one can envision the request to contain anything that defines a set.
%% The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
%% Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.

%% \paragraph {Tasks.}
%% The two main tasks in MFTP are acquiring a file and establishing a relationship between two files.
%% The process of acquiring a file involves sending a request for the file's fragments and receiving fragments sent in response.
%% MFTP is a peer-to-peer protocol because each participant can act as both a client and a server.
%% A peer can send out any fragments for which it gets a request.
%% Fragments are periodically broadcast, even in the absence of requests, to indicate what files exist in the network.
%% A peer can use requests to poll for fragments of a file which can span from individual fragments to the whole file.
%% A peer which has successfully downloaded any number of fragments of a file can then serve those fragments.

%% Matching is a two-step process.
%% The first step applies a matching function $\iota$ to two file IDs to determine whether a match is possible between the two files.  
%% The second step applies a matching function $\mu$ to the two files to determine whether there is a match.
%% If $\mu$($f_1$,$f_2$) is true, $\iota$($fid_1$,$fid_2$) must be true.
%% However, if $\iota$($fid_1$,$fid_2$) is true, $\mu$($f_1$,$f_2$) does \emph{not} have to be true.

@@
